---
title: "GHCNd"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Requirement

+ [R v4.1+](https://www.r-project.org/)

+ [rlang](https://CRAN.R-project.org/package=rlang)

+ [readr](https://CRAN.R-project.org/package=readr)

+ [sf](https://CRAN.R-project.org/package=sf)

+ [dplyr](https://CRAN.R-project.org/package=dplyr)

### Installation

From within R using [remotes](https://CRAN.R-project.org/package=remotes)

```
remotes::install_github("BigelowLab/ghncd")
```

### Usage

#### List the available stations

```{r}
suppressPackageStartupMessages({
  library(rnaturalearth)
  library(rnaturalearthdata)
  library(charlier)
  library(ghcnd)
  library(sf)
  library(ggplot2)
  library(dplyr)
})

stations = fetch_station_list() |>
  glimpse()
```

```{r}
plot(stations['elev'], pch = "+", main = 'Elevation in meters')
```

#### Get Data from one or more stations

```{r}
# stations can come and go, so check that yours are in the current list
my_stations = stations |>
  dplyr::filter(id %in% 
                c('USW00014764', "USW00094623", "USC00172426", "USC00190736"))

x = fetch_station(my_stations) |>
  glimpse()
```
```{r}
x = x |> 
  group_by(NAME)
ggplot(data = x,  aes(x=DATE, y=TMAX)) +
  geom_line() + 
  facet_wrap(~NAME, scales = 'free_y')
```

The same data aggregated into monthly means.  Below, we show January and July TMAX.

```{r}
m = aggregate_monthly(x) |>
  mutate(MONTH = format(DATE, "%b")) |>
  group_by(NAME) 
ggplot(data = filter(m, MONTH == "Jan"),  aes(x=DATE, y=TMAX)) +
  geom_line() + 
  geom_smooth() + 
  facet_wrap(~NAME, scales = 'free_y')
```

```{r}
ggplot(data = filter(m, MONTH == "Jul"),  aes(x=DATE, y=TMAX)) +
  geom_line() + 
  geom_smooth() + 
  facet_wrap(~NAME, scales = 'free_y')
```

And it is always interesting to see how the TMX-TMIN difference changes overtime...

```{r}
m = mutate(m, TDIFF = TMAX-TMIN)

ggplot(data = filter(m, MONTH == "Jul"),  aes(x=DATE, y=TDIFF)) +
  geom_line() + 
  geom_smooth() + 
  facet_wrap(~NAME, scales = 'free_y')
```



### Inventory

GHCN provides an inventory of per station data.   We provide the stations table which allows us to transfer the station names to the inventory.

```{r}
inv = fetch_inventory(stations = stations) |>
  glimpse()
```
